---
output: html_document
editor_options: 
  chunk_output_type: console
---
command & alt & I -> making chunk

---
title: "Enteric virome constrains oral rotavirus vaccine immunogenicity in longitudinally-sampled cohort of Ghanaian infants"
author: "Andrew HyoungJin Kim, George Armah, Francis Dennis, Leran Wang, Rachel Rodgers, Lindsay Droit, Megan T. Baldridge, Scott A. Handley, Vanessa C. Harris"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

**Project Description:** Examination of the 16S rRNA amplicon bacterial microbiome in children from Ghana receiving a three-dose scheduled rotavirus vaccine.

**Collaborator:**
Vanessa Harris (v.harris@aighd.org)

**Other relevant documents:**
The following Phyloseq objects are available. Each is distinguished based on the 16S reference database used for taxonomic classification. RDP and Silva were processed through the species assignment workflow:

* ps0.rotabiome.gg.RDS (GreenGenes)
* ps0.rotabiome.rdp.RDS (RDP)
* ps0.rotabiome.silva.RDS (Silva)

* mapping file: rotabiome_16S_mapping.txt

**References:**
* http://f1000research.com/articles/5-1492/v1

* http://benjjneb.github.io/dada2/tutorial.htm

**Workflow details:** The R commands below represent a full analysis of the following:

1) Sample quality control
2) ASV properties
3) Community Composition
4) Alpha diversity
5) Beta diversity
6) Differential abundance testing

```{r initiate-environment}
# Global knitr option
knitr::opts_chunk$set(fig.width=8,
                      fig.height=6,
                      fig.path="./seroconversion_figures",
                      dev='png',
                      warning=FALSE,
                      message=FALSE)

# Load libraries
library("tidyverse"); packageVersion("tidyverse")
library("phyloseq"); packageVersion("phyloseq")
library("RColorBrewer"); packageVersion("RColorBrewer")
library("vegan"); packageVersion("vegan")
library("knitr"); packageVersion("knitr")
library("ggpubr"); packageVersion("ggpubr")
library("DESeq2"); packageVersion("DESeq2")
library("tidylog"); packageVersion("tidylog")
library("ggplot2"); packageVersion("ggplot2")
library("viridis"); packageVersion("viridis")
library("DAtest"); packageVersion("DAtest")
library("reshape"); packageVersion("reshape")
library("rstatix"); packageVersion("rstatix")
library("remotes"); packageVersion("remotes")
library("phylosmith"); packageVersion("phylosmith")
library("data.table"); packageVersion("data.table")
library("stats"); packageVersion("stats")

```


##Read in data

```{r initiate-data}
# Load Phyloseq Object generated in *_preprocessing.Rmd documents
ps0 <- readRDS(file.choose())
ps0

# Load mapping file
map <- import_qiime_sample_data(file.choose())
ps0 <- merge_phyloseq(ps0, map)
ps0
tax_table(ps0)

# View sample variables
sample_variables(ps0)

# Remove sample that has mostly missing (NA) information
ps0 <- subset_samples(ps0, seq_id != "356.Armah.19888")

# Remove sample with incorrect age data
ps0 <- subset_samples(ps0, seq_id != "179.Armah.19866")

# Remove late collected Dose 1 samples (> 100 days out)
ps0 <- subset_samples(ps0, seq_id != "364.Armah.19896")
ps0 <- subset_samples(ps0, seq_id != "23.Armah.19931")

# Remove late collected Dose 2 sample (> 130 days out)
ps0 <- subset_samples(ps0, seq_id != "141.Armah.20044")

ps0

# View sample variables & generate basic stats
sample_variables(ps0)
sd(sample_sums(ps0))
get_taxa_unique(ps0,"Phylum")
ntaxa(ps0)
nsamples(ps0)
```

##Factor reordering and renaming

```{r factor-adjustments}
# Factor re-ordering, relabelling, etc.
# Reorder it so pre appears before post
levels(sample_data(ps0)$pre_post)
sample_data(ps0)$pre_post <- factor(sample_data(ps0)$pre_post, levels = c("pre", "post"))
levels(sample_data(ps0)$pre_post)

# Convert seroconversion to factors
sample_data(ps0)$seroconversion <- as.factor(sample_data(ps0)$seroconversion)

# Convert factors into numeric
sample_data(ps0)$age_stool_collection <- as.numeric(as.character(sample_data(ps0)$age_stool_collection))

sample_data(ps0)$days_pre_or_post_vaccination <- as.numeric(as.character(sample_data(ps0)$days_pre_or_post_vaccination))

sample_data(ps0)$age_at_first_vac <- as.numeric(as.character(sample_data(ps0)$age_at_first_vac))

sample_data(ps0)$age_at_final_vac <- as.numeric(as.character(sample_data(ps0)$age_at_final_vac))

sample_data(ps0)$age_at_blood_draw <- as.numeric(as.character(sample_data(ps0)$age_at_blood_draw))

sample_data(ps0)$days_between_last_vaccination_and_blood_draw <- as.numeric(as.character(sample_data(ps0)$days_between_last_vaccination_and_blood_draw))

sample_data(ps0)$height_cm <- as.numeric(as.character(sample_data(ps0)$height_cm))

sample_data(ps0)$weight_kg <- as.numeric(as.character(sample_data(ps0)$weight_kg))

sample_data(ps0)$postvaccination_IgA_titer <- as.numeric(as.character(sample_data(ps0)$postvaccination_IgA_titer))

```

##ASV summary statistics

```{r data-assessment}
# Create a new data frame of the sorted row sums, a column of sorted values from 1 to the total number of individuals/counts for each RSV and a categorical variable stating these are all RSVs.
readsumsdf = data.frame(nreads = sort(taxa_sums(ps0), TRUE), 
                        sorted = 1:ntaxa(ps0),
                        type = "ASVs")

# Add a column of sample sums (total number of individuals per sample)
readsumsdf = rbind(readsumsdf,
                   data.frame(nreads = sort(sample_sums(ps0), TRUE),
                              sorted = 1:nsamples(ps0),
                              type = "Samples"))

# Make a data frame with a column for the read counts of each sample for histogram production
sample_sum_df <- data.frame(sum = sample_sums(ps0))

# Make plots
# Generates a bar plot with # of reads (y-axis) for each taxa. Sorted from most to least abundant
# Generates a second bar plot with # of reads (y-axis) per sample. Sorted from most to least
p.reads = ggplot(readsumsdf, aes(x = sorted, y = nreads)) +
  geom_bar(stat = "identity") +
  ggtitle("ASV Assessment") +
  scale_y_log10() +
  facet_wrap(~type, scales = "free") +
  ylab("# of Reads")

# Histogram of the number of Samples (y-axis) at various read depths
p.reads.hist <- ggplot(sample_sum_df, aes(x = sum)) + 
  geom_histogram(color = "black", fill = "firebrick3", binwidth = 150) +
  ggtitle("Distribution of sample sequencing depth") + 
  xlab("Read counts") +
  ylab("# of Samples")

# Final plot, side-by-side
ggarrange(p.reads, p.reads.hist, ncol = 2)

# Basic summary statistics
summary(sample_sums(ps0))
sd(sample_sums(ps0))
```
##Sample assessment

```{r sample-removal-identification}
# Format a data table to combine sample summary data with sample variable data
ss <- sample_sums(ps0)
sd <- as.data.frame(sample_data(ps0)) # useful to coerce the phyloseq object into an R data frame
ss.df <- merge(sd, data.frame("ASVs" = ss), by ="row.names") # merge ss with sd by row names. Rename ss to RSVs in the new data frame

# Plot the data by the treatment variable
y = 1e4 # Set a threshold for the minimum number of acceptable reads. Can start as a guess
x = "seroconversion" # Set the x-axis variable you want to examine

p.ss.boxplot <- ggplot(ss.df, aes(x=age_stool_collection, y = ASVs, color = seroconversion)) +
  geom_point(alpha = 0.6, size = 3, shape = "square") +
  scale_y_log10() +
  geom_hline(yintercept = y, lty = 2) +
  labs(x = "Age @ Stool Collection")
p.ss.boxplot

```

```{r sample-outlier-removal}
# Remove samples with fewer than 1,000 sequences
nsamples(ps0)
ps1 <- prune_samples(sample_sums(ps0) >= y, ps0)
nsamples(ps1)
min(sample_sums(ps1))
```
##Prevalence assessment

```{r prevalence-assessment}
# Begin by removing sequences that were classified as either mitochondria or chlorplast
ps1 # Check the number of taxa prior to removal
ps2 <- ps1 %>%
  subset_taxa(
    Family  != "mitochondria" &
    Class   != "Chloroplast" &
    Phylum != "Cyanobacteria/Chloroplast"
  )
ps2 # Confirm that the taxa were removed

# Prevelance estimation
# Calculate feature prevelance across the data set
prevdf <- apply(X = otu_table(ps2),MARGIN = ifelse(taxa_are_rows(ps2), yes = 1, no = 2),FUN = function(x){sum(x > 0)})

# Add taxonomy and total read counts to prevdf
prevdf <- data.frame(Prevalence = prevdf, TotalAbundance = taxa_sums(ps2), tax_table(ps2))

# Create a table of Phylum, their mean abundances across all samples, and the number of samples they were detected in
plyr::ddply(prevdf, "Phylum", function(df1){cbind(mean(df1$Prevalence),sum(df1$Prevalence))})

# Reorder based on visual inspection of relative dominance of each taxa for plotting

prevdf$Phylum <- factor(prevdf$Phylum, levels = c("Firmicutes", "Actinobacteria", "Proteobacteria", "Bacteroidetes", "Verrucomicrobia", "Fusobacteria", "Deinococcus-Thermus", "Euryarchaeota", "Deferribacteres", "Lentisphaerae", "Spirochaetes", "Synergistetes", "Tenericutes"))

#Prevalence plot
prevdf1 <- subset(prevdf, Phylum %in% get_taxa_unique(ps2, "Phylum"))
p.prevdf1 <- ggplot(prevdf1, aes(TotalAbundance, Prevalence / nsamples(ps2),color=Genus)) +
  geom_hline(yintercept = 0.03, alpha = 0.5, linetype = 2) +
  geom_point(size = 3, alpha = 0.7) +
  scale_x_log10() +
  xlab("Total Abundance") + ylab("Prevalence [Frac. Samples]") +
  facet_wrap(~Phylum) +
  theme(legend.position="none") +
  ggtitle("Phylum Prevelence in All Samples\nColored by Genus")
p.prevdf1

```
Samples are dominated by bacteria from the Firmicutes, Actinobacteria, Proteobacteria and Bacteroidetes with less prevalent members of the Verrumicrobia and Fusobacteria. There are a large number of taxa represented in very few samples. Deinococcus-Therums, Lentispharae, Spirochates, Synergistetes and Teniericutes are very rare appearing in on one sample and can likely be removed or ignored from subsequent analysis.

```{r taxon-cleaning}
# Remove specific taxa
# Define a variable with taxa to remove
get_taxa_unique(ps2, "Phylum")
filterPhyla = c("Lentisphaerae", "Spirochaetes", "Synergistetes", "Deinococcus-Thermus", "Deferribacteres", "Tenericutes", "Euryarchaeota", "Verrucomicrobia", "Fusobacteria")

ps2 # Check the number of taxa prior to removal
ps2 <- subset_taxa(ps2, !Phylum %in% filterPhyla) 
ps2 # Confirm the taxa were removed
get_taxa_unique(ps2, "Phylum")
summary(ps2)
```

# Average OTU counts per subject within dose

```{r Normalization}
##### Normalization #######

# Scales reads by 
# 1) taking proportions,
# 2) multiplying by a given library size of n
# 3) rounding down
scale_reads <- function(physeq, n) {
  physeq.scale <-
    transform_sample_counts(physeq, function(x) {
      (n * x/sum(x))
    })
  otu_table(physeq.scale) <- floor(otu_table(physeq.scale))
  physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
  return(physeq.scale)
}
```


```{r assign color to each phylum}

Phylumcolor<-c("Actinobacteria"="#c51b7d","Bacteroidetes"= "#e9a3c9", "Firmicutes"= "#fde0ef", "Proteobacteria"="#a1d76a")

```


# gut bacterial composition difference in pre- vs. post-vaccination 

```{r pre- vs. post-vaccionation samples @ Dose 1}
# # # samples with pre & post
# # subjectID
# # s47
# # s110
# # s116
# # s149
# # s182
# # s238
# # s442
# # s53
# # s131
# # s248
# # s100
# # s102
# # s59
# # s441
# # s372

prepost_ps <- scale_reads(ps2, 1e5)

levels(sample_data(prepost_ps)$vaccination_proximity)
prepost_ps_DS1 <- subset_samples(prepost_ps, vaccination_proximity == "DS1")
levels(sample_data(prepost_ps_DS1)$vaccination_proximity)

prepost_ps_DS1 <- subset_samples(prepost_ps_DS1, subjectID == "s47" | subjectID == "s110" | subjectID == "s116" | subjectID == "s149" | subjectID == "s182" | subjectID == "s238" | subjectID == "s442" | subjectID == "s53" | subjectID == "s131" | subjectID == "s248" | subjectID == "s100" | subjectID == "s102" | subjectID == "s59" | subjectID == "s441" | subjectID == "372")

prepost_ps_DS1 <- subset_samples(prepost_ps_DS1, stool_number == "1" | stool_number == "2")

# Calculate relative abundance
threshold = 0.01

prepost_ps_DS1_phylum <- prepost_ps_DS1 %>%
  tax_glom(taxrank = "Phylum") %>%                     # agglomerate at class level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > threshold) %>%                    # Filter out low abundance taxa
  arrange(Phylum, Abundance)                           # Sort data frame alphabetically by phylum

# plot relative abundance between pre- & post-vaccination samples @ Dose 1
prepost_phylum_RA <- ggplot(prepost_ps_DS1_phylum, aes(x = as.factor(subjectID), y = Abundance, fill = Phylum)) + theme_bw() +
  geom_bar(stat = "identity", width = 0.85, position="fill") +
  facet_wrap(seroconversion~pre_post, scales = "free_x", nrow=1)+
  labs(title= "", y = "Relative abundance", x = "") +
  theme(legend.position = "bottom",  panel.grid = element_blank()) +
  theme(plot.title = element_text(hjust=0, angle = 0, family= "Arial", size = 12, face="bold"), axis.text = element_text(angle = 90, family= "Arial", size = 12), plot.margin=unit(c(0,0,0,0), "cm"))+
  scale_fill_manual(values = Phylumcolor)
prepost_phylum_RA

# plot relative abundance boxplot of each phyla
prepost_phylum_RA_box <- ggboxplot(prepost_ps_DS1_phylum, x = "pre_post", y = "Abundance", outlier.shape = NA, add = "jitter", color = "pre_post") +
  facet_wrap(seroconversion~Phylum, nrow = 1) +
  stat_compare_means(label = "p.signif", label.x= 2.0, label.y=1.1) +
  ylim(0,1.2) +
  labs(x="", y = "Dose 1\n")+
  theme(plot.title = element_text(hjust=0), axis.text = element_text(angle = 0, family= "Arial", size = 10))+
  theme(panel.grid.major = element_line(), panel.grid.minor = element_line())
prepost_phylum_RA_box

# ggsave("prepost_phylum_RA_box.tiff", units="in", width=8, height=6, dpi=300, compression = 'lzw')

# run statistical test
prepost_wilcox <- prepost_ps_DS1_phylum %>%
group_by(Phylum) %>%
wilcox_test(formula = Abundance~pre_post)
prepost_wilcox

# alpha diversity
# Calculate Alpha Diversity
prepost_alpha_calc <- estimate_richness(prepost_ps_DS1, measures = c("Observed", "Shannon"))

# read merged dataframes
prepost_alpha <- read.delim("prepost_alpha_merged.txt")

levels(prepost_alpha$pre_post)
prepost_alpha$pre_post <- factor(prepost_alpha$pre_post, levels = c("pre", "post"))
levels(prepost_alpha$pre_post)

# paired alpha diversity plots

prepost_rich_paired <- ggpaired(data = prepost_alpha, x = "pre_post", y = "Observed", id = "id", color = "pre_post", point.size = 2.5,line.color = "gray", line.size = 0.4)+ stat_compare_means(method = "wilcox.test", hide.ns = FALSE, label = "p.format", label.x= 2.0, label.y=75) +
  labs(title= "Richness", x= "", y = "") +
   theme(plot.title = element_text(hjust=0.5, angle = 0, family= "Arial", size = 10), legend.position= "none", axis.text = element_text(angle = 0, family= "Arial", size = 10), plot.margin=unit(c(0.2,0,-0.5,0), "cm"))+
  ylim(0,75)

prepost_shannon_paired <- ggpaired(data = prepost_alpha, x = "pre_post", y = "Shannon", id = "id", color = "pre_post", point.size = 2.5,line.color = "gray", line.size = 0.4)+ stat_compare_means(method = "wilcox.test", hide.ns = FALSE, label = "p.format", label.x= 2.0, label.y=4) +
  labs(title= "Shannon", x= "", y = "") +
   theme(plot.title = element_text(hjust=0.5, angle = 0, family= "Arial", size = 10), legend.position= "none", axis.text = element_text(angle = 0, family= "Arial", size = 10), plot.margin=unit(c(0.2,0,-0.5,0), "cm"))+
  ylim(0,4)

prepost_alpha_fig <- ggarrange(prepost_rich_paired, prepost_shannon_paired, ncol=2 )
prepost_alpha_fig

ggsave("prepost_alpha_fig.tiff", units="in", width=8, height=3, dpi=300, compression = 'lzw')


# beta diversity

ord.psm.prepost.wuni <- ordinate(prepost_ps_DS1, method = "PCoA", distance = "wunifrac")
evals.wuni.prepost <- ord.psm.prepost.wuni$values$Eigenvalues #Update

# PCoA plots
p.ord.wuni.prepost <- plot_ordination(prepost_ps_DS1, ord.psm.prepost.wuni, color = "pre_post") + geom_point(size = 2.5)+
  labs(title= "", color = "pre_post")+
   theme(plot.title = element_text(hjust=0.5, angle = 0, family= "Arial", size = 10), legend.position= "bottom", axis.text = element_text(angle = 0, family= "Arial", size = 10))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
    geom_hline(yintercept = 0, alpha = 0.5, linetype = 2) +
   geom_vline(xintercept = 0, alpha = 0.5, linetype = 2)
p.ord.wuni.prepost

ggsave("p.ord.wuni.prepost.tiff", units="in", width=8, height=3, dpi=300, compression = 'lzw')

# run adonis

pruned.prepost <- prune_taxa(taxa_sums(prepost_ps_DS1) > 0, prepost_ps_DS1)
pruned.prepost <- prune_samples(sample_sums(pruned.prepost) > 0, prepost_ps_DS1)

sd.prepost <- data.frame(sample_data(pruned.prepost))

# Adonis test
ds1.bray <- phyloseq::distance(pruned.prepost, method = "wunifrac")
vegan::adonis2(ds1.bray ~ pre_post, data = sd.prepost)

```


```{r average-ASV-per-subject-per-dose}
###### Merge functions ############

#Merge samples by averaging OTU countsinstead of summing
merge_samples_mean <- function(physeq, group){
  # Calculate the number of samples in each group
  group_sums <- as.matrix(table(sample_data(physeq)[ ,group]))[,1]
  
  # Merge samples by summing
  merged <- merge_samples(physeq, group)
  
  # Divide summed OTU counts by number of samples in each group to get mean
  # Calculation is done while taxa are columns
  x <- as.matrix(otu_table(merged))
  if(taxa_are_rows(merged)){ x<-t(x) }
  out <- floor(t(x/group_sums))
  
  # Return new phyloseq object with
  
  out <- otu_table(out, taxa_are_rows = TRUE)
  otu_table(merged) <- out
  return(merged)
}

# Sacle Reads
ps3 <- scale_reads(ps2, 1e5)

# Subset by dose
levels(sample_data(ps3)$vaccination_proximity)
ps3.DS1 <- subset_samples(ps3, vaccination_proximity == "DS1")
levels(sample_data(ps3.DS1)$vaccination_proximity)
summary(taxa_sums(ps3.DS1))
ps3.DS1 <- prune_taxa(taxa_sums(ps3.DS1)>0, ps3.DS1)
summary(taxa_sums(ps3.DS1))

ps3.DS2 <- subset_samples(ps3, vaccination_proximity == "DS2")
levels(sample_data(ps3.DS2)$vaccination_proximity)
summary(taxa_sums(ps3.DS2))
ps3.DS2 <- prune_taxa(taxa_sums(ps3.DS2)>0, ps3.DS2)
summary(taxa_sums(ps3.DS2))

ps3.DS3 <- subset_samples(ps3, vaccination_proximity == "DS3")
levels(sample_data(ps3.DS3)$vaccination_proximity)
summary(taxa_sums(ps3.DS3))
ps3.DS3 <- prune_taxa(taxa_sums(ps3.DS3)>0, ps3.DS3)
summary(taxa_sums(ps3.DS3))

# Count number of subjects
length(unique(sample_data(ps3)$subjectID))
length(unique(sample_data(ps3.DS1)$subjectID))
length(unique(sample_data(ps3.DS2)$subjectID))
length(unique(sample_data(ps3.DS3)$subjectID))

# Merge samples on subjectID
psm <- merge_samples_mean(ps3, "subjectID")
nsamples(psm)
length(unique(sample_data(ps3)$subjectID))

psm.DS1 <- merge_samples_mean(ps3.DS1, "subjectID")
nsamples(psm.DS1)
length(unique(sample_data(ps3.DS1)$subjectID))

psm.DS2 <- merge_samples_mean(ps3.DS2, "subjectID")
nsamples(psm.DS2)
length(unique(sample_data(ps3.DS2)$subjectID))

psm.DS3 <- merge_samples_mean(ps3.DS3, "subjectID")
nsamples(psm.DS3)
length(unique(sample_data(ps3.DS3)$subjectID))

# 'fix' subject ids by renaming merged subjectID with rownames
# 'fix' seroconversion
sample_data(psm.DS1)$subjectID
sample_data(psm.DS1)$subjectID <- rownames(sample_data(psm.DS1))
sample_data(psm.DS1)$subjectID

sample_data(psm.DS1)$seroconversion
sample_data(psm.DS1)$seroconversion <-  factor(sample_data(psm.DS1)$seroconversion, labels = c("No","Yes"))
sample_data(psm.DS1)$seroconversion

sample_data(psm.DS1)$vaccination_proximity
sample_data(psm.DS1)$vaccination_proximity <- "DS1"
sample_data(psm.DS1)$vaccination_proximity

sample_data(psm.DS2)$subjectID
sample_data(psm.DS2)$subjectID <- rownames(sample_data(psm.DS2))
sample_data(psm.DS2)$seroconversion
sample_data(psm.DS2)$seroconversion <- factor(sample_data(psm.DS2)$seroconversion, labels = c("No", "Yes"))
sample_data(psm.DS2)$vaccination_proximity
sample_data(psm.DS2)$vaccination_proximity <- "DS2"
sample_data(psm.DS2)$subjectID
sample_data(psm.DS2)$seroconversion
sample_data(psm.DS2)$vaccination_proximity

sample_data(psm.DS3)$subjectID
sample_data(psm.DS3)$subjectID <- rownames(sample_data(psm.DS3))
sample_data(psm.DS3)$seroconversion
sample_data(psm.DS3)$seroconversion <- factor(sample_data(psm.DS3)$seroconversion, labels = c("No", "Yes"))
sample_data(psm.DS3)$vaccination_proximity
sample_data(psm.DS3)$vaccination_proximity <- "DS3"
sample_data(psm.DS3)$subjectID
sample_data(psm.DS3)$seroconversion
sample_data(psm.DS3)$vaccination_proximity
```

```{r IgA distribution @ Dose 1}
averaged_raw <- psm.DS1 %>%
psmelt() %>%
mutate(postvaccination_IgA_titer = as.numeric(postvaccination_IgA_titer)) %>%
arrange(postvaccination_IgA_titer)

#get the correct order
new.order <- psm.DS1 %>%
psmelt() %>%
arrange(postvaccination_IgA_titer) %>%
distinct(subjectID) %>%
deframe()

averaged_raw$subjectID <- factor(averaged_raw$subjectID,levels = new.order)

IgA_distri <- ggplot(averaged_raw, aes(x = subjectID, y = postvaccination_IgA_titer, fill=seroconversion))+
theme_bw() +
geom_bar(stat = "identity")+
theme(axis.text.x = element_text(angle = 90))

IgA_distri


IgA_box <- ggplot(averaged_raw, aes(x = seroconversion, y = postvaccination_IgA_titer, fill=seroconversion))+
theme_bw() +
geom_boxplot()+
geom_point()

IgA_box
```

#Figure_1_Relative_abundace_Phylum
```{r community-composition-plots}
# Community composition
# Phyla level plots
# Melt to long format (for ggploting) 
# Prune out phyla below % in each sample
threshold = 0.01

# Dose 1
# Counts Averaged Per Subject
psm.DS1_phylum <- psm.DS1 %>%
  tax_glom(taxrank = "Phylum") %>%                     # agglomerate at class level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > threshold) %>%                    # Filter out low abundance taxa
  arrange(Phylum, Abundance)                           # Sort data frame alphabetically by phylum

# write.table(psm.DS1_Genus, file = "psm.DS1_genus.txt", sep = "\t",
#             row.names = TRUE, col.names = NA)

# Dose 1
# Counts Averaged Per Subject
p.comm.avg.DS1 <- ggplot(psm.DS1_phylum, aes(x = as.factor(seroconversion), y = Abundance, fill = Phylum)) + theme_bw() +
  geom_bar(stat = "identity", width = 0.85, position="fill") +
  facet_grid(~vaccination_proximity, scales = "free", space = "free") +
  facet_wrap(~pre_post, nrow=1)+
  labs(title= "", y = "Relative abundance", x = "") +
  theme(legend.position = "",  panel.grid = element_blank()) +
  theme(plot.title = element_text(hjust=0, angle = 0, family= "Arial", size = 12, face="bold"), axis.text = element_text(angle = 0, family= "Arial", size = 12), plot.margin=unit(c(0,0,0,0), "cm"))+
  scale_fill_manual(values = Phylumcolor)
p.comm.avg.DS1

# Dose 2
# Counts Averaged Per Subject
psm.DS2_phylum <- psm.DS2 %>%
  tax_glom(taxrank = "Phylum") %>%                     # agglomerate at class level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > threshold) %>%                         # Filter out low abundance taxa
  arrange(Phylum, Abundance)                           # Sort data frame alphabetically by phylum

p.comm.avg.DS2 <- ggplot(psm.DS2_phylum, aes(x = as.factor(seroconversion), y = Abundance, fill = Phylum)) + theme_bw() +
  geom_bar(stat = "identity", width = 0.85, position="fill") +
  facet_grid(~vaccination_proximity, scales = "free", space = "free") +
  labs(title= "", y = "", x = "seroconversion") +
  theme(legend.position = "",  panel.grid = element_blank()) +
  theme(plot.title = element_text(hjust=0, angle = 0, family= "Arial", size = 12, face="bold"), axis.text = element_text(angle = 0, family= "Arial", size = 12), plot.margin=unit(c(0,0,0,0), "cm"), axis.text.y = element_blank(), axis.ticks.y = element_blank())+
  scale_fill_manual(values = Phylumcolor)

p.comm.avg.DS2

# Dose 3
# Counts Averaged Per Subject
psm.DS3_phylum <- psm.DS3 %>%
  tax_glom(taxrank = "Phylum") %>%                     # agglomerate at class level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > threshold) %>%                         # Filter out low abundance taxa
  arrange(Phylum, Abundance)                           # Sort data frame alphabetically by phylum

p.comm.avg.DS3 <- ggplot(psm.DS3_phylum, aes(x = as.factor(seroconversion), y = Abundance, fill = Phylum)) + theme_bw() +
  geom_bar(stat = "identity", width = 0.85, position="fill") +
  facet_grid(~vaccination_proximity, scales = "free", space = "free") +
  labs(title= "", y = "", x = "") +
  theme(legend.position = "",  panel.grid = element_blank()) +
  theme(plot.title = element_text(hjust=0, angle = 0, family= "Arial", size = 12, face="bold"), axis.text = element_text(angle = 0, family= "Arial", size = 12), plot.margin=unit(c(0,0,0,0), "cm"), axis.text.y = element_blank(), axis.ticks.y = element_blank())+
  scale_fill_manual(values = Phylumcolor)

p.comm.avg.DS3

Figure2A<-ggarrange(p.comm.avg.DS1, p.comm.avg.DS2, p.comm.avg.DS3, ncol = 3, legend = "bottom", common.legend = TRUE, widths = c(1.1,1,1))

# ggsave("Rotabiome_bacteriome_seroconversion_Figure2A.tiff", units="in", width=8, height=6, dpi=300, compression = 'lzw')

```

#supfigure1_class_level_boxplots
```{r phyla-level-boxplots}

# Transform each averaged dose to relative abundance
psm.ds1.ra <- transform_sample_counts(psm.DS1, function(x) {x/sum(x)} )
psm.ds2.ra <- transform_sample_counts(psm.DS2, function(x) {x/sum(x)} )
psm.ds3.ra <- transform_sample_counts(psm.DS3, function(x) {x/sum(x)} )

# Agglomerate average counts at each dose
glom.ds1 <- tax_glom(psm.ds1.ra, taxrank = "Phylum")
glom.ds2 <- tax_glom(psm.ds2.ra, taxrank = "Phylum")
glom.ds3 <- tax_glom(psm.ds3.ra, taxrank = "Phylum")

# Create dataframe of each
dat.ds1 <- as.tibble(psmelt(glom.ds1))
dat.ds2 <- as.tibble(psmelt(glom.ds2))
dat.ds3 <- as.tibble(psmelt(glom.ds3))

# Reorder
# Filter low abundance taxa
# Filtration of nsamples by removing nsamples <1

levels(dat.ds1$Phylum)
dat.ds1.reorder <- dat.ds1 %>%
  filter(Abundance > threshold) %>%
  droplevels()
levels(dat.ds1.reorder$Phylum)


levels(dat.ds2$Phylum)
dat.ds2.reorder <- dat.ds2 %>%
  filter(Abundance > threshold) %>%
  droplevels()
levels(dat.ds2.reorder$Phylum)

levels(dat.ds3$Phylum)
dat.ds3.reorder <- dat.ds3 %>%
  filter(Abundance > threshold) %>%
  droplevels() 
levels(dat.ds3.reorder$Phylum)

# Boxplots
p.box.Phylum.ds1 <- ggboxplot(dat.ds1.reorder, x = "seroconversion", y = "Abundance", outlier.shape = NA, add = "jitter", color = "seroconversion") +
  facet_wrap(~Phylum, nrow = 1) +
  stat_compare_means(label = "p.signif", label.x= 2.0, label.y=1.1) +
  ylim(0,1.2) +
  labs(x="", y = "Dose 1\n")+
  theme(plot.title = element_text(hjust=0), axis.text = element_text(angle = 0, family= "Arial", size = 10))+
  theme(panel.grid.major = element_line(), panel.grid.minor = element_line())

p.box.Phylum.ds2 <- ggboxplot(dat.ds2.reorder, x = "seroconversion", y = "Abundance", outlier.shape = NA, add = "jitter", color = "seroconversion") +
  facet_wrap(~Phylum, nrow = 1) +
  stat_compare_means(label = "p.signif", label.x= 2.0, label.y=1.1) +
  ylim(0,1.2) +
  labs(x="", y = "Dose 2\nRelative abundance")+
  theme(plot.title = element_text(hjust=0), axis.text = element_text(angle = 0, family= "Arial", size = 10))+
  theme(panel.grid.major = element_line(), panel.grid.minor = element_line())


p.box.Phylum.ds3 <- ggboxplot(dat.ds3.reorder, x = "seroconversion", y = "Abundance", outlier.shape = NA, add = "jitter", color = "seroconversion") +
  facet_wrap(~Phylum, nrow = 1) +
  stat_compare_means(label = "p.signif", label.x= 2.0, label.y=1.1) +
  ylim(0,1.2) +
  labs(x="", y="Dose 3\n")+
  theme(plot.title = element_text(hjust=0), axis.text = element_text(angle = 0, family= "Arial", size = 10))+
  theme(panel.grid.major = element_line(), panel.grid.minor = element_line())


FigureS1A<-ggarrange(p.box.Phylum.ds1, p.box.Phylum.ds2, p.box.Phylum.ds3, nrow = 3,common.legend = TRUE, legend = "bottom")

# ggsave("Rotabiome_bacteriome_FigureS1A.tiff", units="in", width=8, height=6, dpi=300, compression = 'lzw')

FigureS1.all.df <- bind_rows(dat.ds1.reorder, dat.ds2.reorder, dat.ds3.reorder)

p.box.Phylum.all <- ggplot(FigureS1.all.df, aes(x = vaccination_proximity, y= Abundance, group = Phylum, color = Phylum)) +
  geom_jitter(width = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~seroconversion,nrow = 1) +
  labs(y = "Relative abundance", x = "", color = "Phylum")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")

# ggsave("Rotabiome_bacteriome_FigureS1B.tiff", units="in", width=8, height=6, dpi=300, compression = 'lzw')


# Statistical analysis for FigureS1B
  
FigureS1.all.df.aov <- aov(Abundance ~ vaccination_proximity + seroconversion, data = FigureS1.all.df)
summary(FigureS1.all.df.aov)

```


## alpha diversity
```{r add-sample-data}

# Calculate Alpha Diversity
div.DS1 <- estimate_richness(psm.DS1, measures = c("Observed", "Shannon"))
div.DS2 <- estimate_richness(psm.DS2, measures = c("Observed", "Shannon"))
div.DS3 <- estimate_richness(psm.DS3, measures = c("Observed", "Shannon"))

# Merge with sample data
sd.psm.DS1 <- as.data.frame(sample_data(psm.DS1))
sd.psm.DS2 <- as.data.frame(sample_data(psm.DS2))
sd.psm.DS3 <- as.data.frame(sample_data(psm.DS3))

psm.DS1.rich <- merge(sd.psm.DS1, div.DS1, by = "row.names")
psm.DS2.rich <- merge(sd.psm.DS2, div.DS2, by = "row.names")
psm.DS3.rich <- merge(sd.psm.DS3, div.DS3, by = "row.names")

# Bind all 3 vaccination proximity alpha diversity values into a single data frame
bac.alpha.all.df <- bind_rows(psm.DS1.rich, psm.DS2.rich, psm.DS3.rich)

# Add a continuous time variable
bac.alpha.timeseries.df <- bac.alpha.all.df %>%
  as_tibble %>%
  mutate(time_point = ifelse(vaccination_proximity == "DS1", 1,
                             ifelse(vaccination_proximity == "DS2", 2, 3)))

# Test for linear association between time and Observed
fit <- lm(Observed ~ time_point, data = bac.alpha.timeseries.df)
summary(fit)

# Plot
ggplot(bac.alpha.timeseries.df, aes(x=vaccination_proximity, y=Observed, group = seroconversion, color = seroconversion)) +
  geom_jitter(width = 0.2) +
  geom_smooth(method = "lm") +
  labs(y = "Richness", x = "", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "none")+
  ylim(0,150)

# ggsave("Rotabiome_phage_seroconversion_Figure2B_richness.tiff", units="in", width=4, height=4, dpi=300, compression = 'lzw')

fit.2 <- aov(Observed ~ time_point + seroconversion, data = bac.alpha.timeseries.df)
summary(fit.2)


# Test for linear association between time and Observed
fit <- lm(Observed ~ time_point, data = bac.alpha.timeseries.df)
summary(fit)

# Plot
ggplot(bac.alpha.timeseries.df, aes(x=vaccination_proximity, y=Shannon, group = seroconversion, color = seroconversion)) +
  geom_jitter(width = 0.2) +
  geom_smooth(method = "lm") +
  labs(y = "Shannon diversity", x = "", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "none")+
  ylim(0,4)

# ggsave("Rotabiome_phage_seroconversion_Figure2B_diversity.tiff", units="in", width=4, height=4, dpi=300, compression = 'lzw')


fit.2 <- aov(Shannon ~ time_point + seroconversion, data = bac.alpha.timeseries.df)
summary(fit.2)

```

# Beta diversity

```{r ordination-on-averaged}
# Dose 1
# Calculate distances
ord.psm.DS1.wuni <- ordinate(psm.DS1, method = "PCoA", distance = "wunifrac")
evals.wuni.DS1 <- ord.psm.DS1.wuni$values$Eigenvalues #Update

# PCoA plots
p.ord.wuni.DS1 <- plot_ordination(psm.DS1, ord.psm.DS1.wuni, color = "seroconversion") +
labs(title= "", color = "seroconversion")+
   theme(plot.title = element_text(hjust=0.5, angle = 0, family= "Arial", size = 10), legend.position= "none", axis.text = element_text(angle = 0, family= "Arial", size = 10))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
    geom_hline(yintercept = 0, alpha = 0.5, linetype = 2) +
   geom_vline(xintercept = 0, alpha = 0.5, linetype = 2) 


# Dose 2
# Calculate distances
ord.psm.DS2.wuni <- ordinate(psm.DS2, method = "PCoA", distance = "wunifrac")
evals.wuni.DS2 <- ord.psm.DS2.wuni$values$Eigenvalues #Update


# PCoA plots
p.ord.wuni.DS2 <- plot_ordination(psm.DS2, ord.psm.DS2.wuni, color = "seroconversion") +
  geom_point(size=2, alpha = 0.7) +
 labs(color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5, angle = 0, family= "Arial", size = 10), legend.position= "none", axis.text = element_text(angle = 0, family= "Arial", size = 10))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
      geom_hline(yintercept = 0, alpha = 0.5, linetype = 2) +
   geom_vline(xintercept = 0, alpha = 0.5, linetype = 2)


# Dose 3
# Calculate distances
ord.psm.DS3.wuni <- ordinate(psm.DS3, method = "PCoA", distance = "wunifrac")
evals.wuni.DS3 <- ord.psm.DS3.wuni$values$Eigenvalues #Update


# PCoA plots
p.ord.wuni.DS3 <- plot_ordination(psm.DS3, ord.psm.DS3.wuni, color = "seroconversion") +
  geom_point(size=2, alpha = 0.7) +
  labs(color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5, angle = 0, family= "Arial", size = 10), legend.position= "none", axis.text = element_text(angle = 0, family= "Arial", size = 10))+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))+
      geom_hline(yintercept = 0, alpha = 0.5, linetype = 2) +
   geom_vline(xintercept = 0, alpha = 0.5, linetype = 2) 



# ggsave("Rotabiome_bacteriome_seroconversion_Figure2C.tiff", units="in", width=16, height=6, dpi=300, compression = 'lzw')

```

```{r adonis}
## Dose 1
# Remove zero'd taxa
pruned.ds1 <- prune_taxa(taxa_sums(psm.DS1) > 0, psm.DS1)
pruned.ds1 <- prune_samples(sample_sums(pruned.ds1) > 0, psm.DS1)

sd.ds1 <- data.frame(sample_data(pruned.ds1))

# Adonis test
ds1.bray <- phyloseq::distance(pruned.ds1, method = "wunifrac")
vegan::adonis2(ds1.bray ~ seroconversion, data = sd.ds1)

## Dose 2
# Remove zero'd taxa
pruned.ds2 <- prune_taxa(taxa_sums(psm.DS2) > 0, psm.DS2)
pruned.ds2 <- prune_samples(sample_sums(pruned.ds2) > 0, psm.DS2)

sd.ds2 <- data.frame(sample_data(pruned.ds2))

# Adonis test
ds2.bray <- phyloseq::distance(pruned.ds2, method = "wunifrac")
vegan::adonis2(ds2.bray ~ seroconversion, data = sd.ds2)

## Dose 3
# Remove zero'd taxa
pruned.ds3 <- prune_taxa(taxa_sums(psm.DS3) > 0, psm.DS3)
pruned.ds3 <- prune_samples(sample_sums(psm.DS3) > 0, psm.DS3)

sd.ds3 <- data.frame(sample_data(pruned.ds3))

# Adonis test
ds3.bray <- phyloseq::distance(pruned.ds3, method = "wunifrac")
vegan::adonis2(ds3.bray ~ seroconversion, data = sd.ds3)

```

# Plotvolcano function
```{r Plotvolcano}

plotVolcano <- function(ps0,ps0.subset,factorOfInterest = NULL,truthFacet = NULL,alpha = NULL,baseMean = NULL,outputPath = NULL){
  replace_counts = function(physeq, dds) {
    dds_counts = counts(dds, normalized = TRUE)
    if (!identical(taxa_names(physeq), rownames(dds_counts))) {
      stop("OTU ids don’t match")
    }
    otu_table(physeq) = otu_table(dds_counts, taxa_are_rows = TRUE)
    return(physeq)
  }
  formula <- as.formula(paste0("~",factorOfInterest))
  # Make deseq ready object
  ds.all <- phyloseq_to_deseq2(ps0, formula)
  dds.all <- DESeq(ds.all,test="Wald", fitType="local", betaPrior = FALSE)
  # replace the count table in physeq to the normalized count table
  rlog.all <- replace_counts(ps0, dds.all)
  # get a melted table with all the information in it
  rlog.all <- psmelt(rlog.all)
  # rename the "OTU" count to "ASV" column
  rlog.all <- dplyr::rename(rlog.all, "ASV" = "OTU")
  #make deseq ready formula
  formula <- as.formula(paste0("~",factorOfInterest))
  # convert phyloseq to deseq2 table
  ds.subset <- phyloseq_to_deseq2(ps0.subset,formula)
  # run deseq
  dds.subset <- DESeq(ds.subset,test="Wald", fitType="local", betaPrior = FALSE)
  #get deseq result
  res.dds.subset <- results(dds.subset,cooksCutoff = FALSE)
  # get a overview of the deseq result
  summary(res.dds.subset)
  print(mcols(res.dds.subset))  #INFECTED YES vs NO
  # make a dataframe on the deseq result
  res.dds.subset <- as.data.frame(res.dds.subset[ which(res.dds.subset$baseMean >  100), ]) %>%
    rownames_to_column("OTU")
  # make a dataframe on the taxonomy table
  taxa <- data.frame(tax_table(ps0.subset)) %>%
    rownames_to_column("OTU")
  taxa$id <- c(1:dim(taxa)[1])
  taxa$ASVname <- paste0("ASV_",taxa$id)
  # combine deseq result and taxonomy table
  res.test <- left_join(res.dds.subset,taxa,by = "OTU")
  # get a volcano-plot ready table
  cleaned.res.test <- res.test %>%
    mutate(significant = res.test$padj < alpha)
  outputTable <-  res.test %>%
                  subset(padj < alpha)
  # save table
  write.csv(outputTable,outputPath)
  View(outputTable)
  l <- unique(as.factor(cleaned.res.test$Phylum))
  color <- colorRampPalette(brewer.pal( 9 , "Set1" ))(length(l))
  names(color) <- l
  print(color)
  # valcano plot
  p.volcano <- ggplot(data = cleaned.res.test,
                      aes(x = log2FoldChange,
                          y = -log10(pvalue),
                          color = Phylum,
                          label1 = Family,
                          label2 = Genus,
                          label3 = Phylum)) +
    scale_color_manual(values=Phylumcolor, na.translate=FALSE)+
    geom_point(data = subset(cleaned.res.test,cleaned.res.test$significant == FALSE),
               color = "grey") +
    geom_point(data = subset(cleaned.res.test,cleaned.res.test$significant == TRUE),
               aes(color = Phylum,
                   size = baseMean))+
    geom_vline(xintercept = 0,lty = 2)+
    geom_hline(yintercept = -log10(0.05))+
    geom_text_repel(data = subset(cleaned.res.test,cleaned.res.test$significant == TRUE)
                    ,aes(label= ASVname),show.legend = FALSE) +
    theme_bw() +
    theme(axis.title = element_text(size=12),
          axis.text = element_text(size=12),
          axis.text.x = element_text(hjust = 0, vjust=0.5),
          legend.text = element_text(size=8),
          legend.position = "right")
  print("start to prepare data for ground truth plot:")
  if (dim(outputTable)[1] != 0){
  # ground truth
  # Create merge taxa table
  tax <- as.data.frame(tax_table(ps0))
  tax <-  as.tibble(rownames_to_column(tax, var = "ASV"))
  tax <- left_join(tax,taxa[,c("OTU","ASVname")],by = c("ASV" = "OTU"))
  # subset deseq result data to significant taxa only
  df.subset <- as.data.frame(res.dds.subset[which(res.dds.subset$padj < alpha & res.dds.subset$baseMean>baseMean), ]) %>%                 dplyr::rename("ASV" = "OTU") %>%
               left_join(tax, by = "ASV")
  #plot
  theme_set(theme_bw(base_size =8,
                     base_family = "Arial"))
  df.subset2 <- inner_join(df.subset, rlog.all, by = "ASV")
  formula.truth <- as.formula(paste0(truthFacet,"~ASVname"))
  print("start to plot ground truth plot2:")
  # Boxplots
  p.boxplot.subset2 <- ggboxplot(df.subset2, x=factorOfInterest, y="Abundance", color = "Phylum.y",outlier.shape = NA) +
     scale_color_manual(values=Phylumcolor, na.translate=FALSE)+
    geom_jitter(aes(x = df.subset2[,factorOfInterest], colour = Phylum.y),alpha = 0.7, width = 0.2, size = 1)+
    scale_y_log10() +
    facet_wrap(formula.truth)+
    # labs(title = "Absolute Abundant of Differentially Abundant Taxa") +
    theme(axis.text.x = element_text(hjust = 1)) +
    stat_compare_means(label = "p.signif", hide.ns = TRUE) +
    labs(color='Phylum')
  ggarrange(p.volcano,p.boxplot.subset2)
  }
  else{
    print("There is no significant taxa in this cohort")
    p.volcano
  }
}
```


# Diff Ab: DS1. Averaged Counts
```{r diffab-avg-DS1}
# Filter low abundant taxa
min.DS1 = round(nsamples(psm.DS1) * 0.10)

# Remove low abundant features
DS1.filt <- preDA(psm.DS1, min.samples = min.DS1, min.reads = 100)

DS1<-DS1.filt

plotVolcano(DS1, DS1,"seroconversion","Genus.y", 0.05, 0, "./DS1_seroconversion_DESeq.csv")

```


# Diff Ab: DS2. Averaged Counts

```{r diffab-avg-DS2}
# Filter low abundant taxa
min.DS2 = round(nsamples(psm.DS2) * 0.10)

# Remove low abundant features
DS2.filt <- preDA(psm.DS2, min.samples = min.DS2, min.reads = 100)

DS2 <- DS2.filt 

plotVolcano(DS2, DS2,"seroconversion","Genus.y", 0.05, 0, "./DS2_seroconversion_DESeq.csv")

```

# Diff Ab: DS3. Averaged Counts

```{r diffab-avg-DS3}

# Filter low abundant taxa
min.DS3 = round(nsamples(psm.DS3) * 0.10)

# Remove low abundant features
DS3.filt <- preDA(psm.DS3, min.samples = min.DS3, min.reads = 100)

DS3 <- DS3.filt 

plotVolcano(DS3, DS3,"seroconversion","Genus.y", 0.05, 0, "./DS3_seroconversion_DESeq.csv")

```

```{r ground truth plots from DESeq2 results}
# Filter low abundant taxa
min.DS1 = round(nsamples(psm.DS1) * 0.10)
min.DS2 = round(nsamples(psm.DS2) * 0.10)
min.DS3 = round(nsamples(psm.DS3) * 0.10)
DS1.filt <- preDA(psm.DS1, min.samples = min.DS1, min.reads = 100)
DS2.filt <- preDA(psm.DS2, min.samples = min.DS2, min.reads = 100)
DS3.filt <- preDA(psm.DS3, min.samples = min.DS3, min.reads = 100)
# DS1.filt<- subset_samples(psm.DS1, postvaccination_IgA_titer > "7")
# DS2.filt<- subset_samples(psm.DS2, postvaccination_IgA_titer > "7")
# DS3.filt<- subset_samples(psm.DS3, postvaccination_IgA_titer > "7")

# Dose 1

DS1.filt.cont <- DS1.filt %>%
transform_sample_counts(function(x) {x/sum(x)}) %>%
psmelt()

DS1.Enterococcus <- DS1.filt.cont %>% filter(OTU == "GCAAGCGTTGTCCGGATTTATTGGGCGTAAAGCGAGCGCAGGCGGTTTCTTAAGTCTGATGTGAAAGCCCCCGGCTCAACCGGGGAGGGTCATTGGAAACTGGGAGACTTGAGTGCAGAAGAGGAGAGTGGAATTCCATGTGTAGCGGTGAAATGCGTAGATATATGGAGGAACACCAGTGGCGAAGGCGGCTCTCTGGTCTGTAACTGACGCTG")

DS1.Enterococcus.plot <-ggplot(DS1.Enterococcus, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "DS1.Enterococcus abundance", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")

DS1.Lactobacillus1 <- DS1.filt.cont %>% filter(OTU == "GCAAGCGTTATCCGGATTTATTGGGCGTAAAGCGAGCGCAGGCGGTTTTTTAAGTCTGATGTGAAAGCCCTCGGCTTAACCGAGGAAGCGCATCGGAAACTGGGAAACTTGAGTGCAGAAGAGGACAGTGGAACTCCATGTGTAGCGGTGAAATGCGTAGATATATGGAAGAACACCAGTGGCGAAGGCGGCTGTCTGGTCTGTAACTGACGCTG")

DS1.Lactobacillus1.plot <-ggplot(DS1.Lactobacillus1, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "DS1.Lactobacillus1 abundance", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")

DS1.Lactobacillus2 <- DS1.filt.cont %>% filter(OTU == "GCAAGCGTTGTCCGGATTTATTGGGCGTAAAGCGAGTGCAGGCGGTTCAATAAGTCTGATGTGAAAGCCTTCGGCTCAACCGGAGAATTGCATCAGAAACTGTTGAACTTGAGTGCAGAAGAGGAGAGTGGAACTCCATGTGTAGCGGTGGAATGCGTAGATATATGGAAGAACACCAGTGGCGAAGGCGGCTCTCTGGTCTGCAACTGACGCTG")

DS1.Lactobacillus2.plot <-ggplot(DS1.Lactobacillus2, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "DS1.Lactobacillus2 abundance", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")

DS1.Clostridium_sensu_stricto <- DS1.filt.cont %>% filter(OTU == "GCGAGCGTTATCCGGATTTACTGGGCGTAAAGGGAGCGTAGGCGGATGATTAAGTGGGATGTGAAATACCCGGGCTCAACTTGGGTGCTGCATTCCAAACTGGTTATCTAGAGTGCAGGAGAGGAGAGTGGAATTCCTAGTGTAGCGGTGAAATGCGTAGAGATTAGGAAGAACACCAGTGGCGAAGGCGACTCTCTGGACTGTAACTGACGCTG")

DS1.Clostridium_sensu_stricto.plot <-ggplot(DS1.Clostridium_sensu_stricto, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "DS1.Clostridium_sensu_stricto abundance", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")

DS1.difficile <- DS1.filt.cont %>% filter(OTU == "GCTAGCGTTATCCGGATTTACTGGGCGTAAAGGGTGCGTAGGCGGTCTTTCAAGTCAGGAGTGAAAGGCTACGGCTCAACCGTAGTAAGCTCTTGAAACTGGGAGACTTGAGTGCAGGAGAGGAGAGTGGAATTCCTAGTGTAGCGGTGAAATGCGTAGATATTAGGAGGAACACCAGTTGCGAAGGCGGCTCTCTGGACTGTAACTGACGCTGA")

DS1.difficile.plot <-ggplot(DS1.difficile, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "DS1.difficile abundance", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")


DESeq2_cont_DS1 <-ggarrange(DS1.Enterococcus.plot, DS1.Lactobacillus1.plot, DS1.Lactobacillus2.plot, DS1.Clostridium_sensu_stricto.plot, DS1.difficile.plot, ncol = 3, nrow = 2)

# Dose 2

DS2.filt.cont <- DS2.filt %>%
transform_sample_counts(function(x) {x/sum(x)}) %>%
psmelt()

DS2.Staphylococcus <- DS2.filt.cont %>% filter(OTU == "GCAAGCGTTATCCGGAATTATTGGGCGTAAAGCGCGCGTAGGCGGTTTTTTAAGTCTGATGTGAAAGCCCACGGCTCAACCGTGGAGGGTCATTGGAAACTGGAAAACTTGAGTGCAGAAGAGGAAAGTGGAATTCCATGTGTAGCGGTGAAATGCGCAGAGATATGGAGGAACACCAGTGGCGAAGGCGACTTTCTGGTCTGTAACTGACGCTG")

DS2.Staphylococcus.plot <-ggplot(DS2.Staphylococcus, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "DS2.Staphylococcus abundance", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")

DS2.Proteus <- DS2.filt.cont %>% filter(OTU == "GCAAGCGTTAATCGGAATTACTGGGCGTAAAGCGCACGCAGGCGGTCAATTAAGTCAGATGTGAAAGCCCCGAGCTTAACTTGGGAATTGCATCTGAAACTGGTTGGCTAGAGTCTTGTAGAGGGGGGTAGAATTCCATGTGTAGCGGTGAAATGCGTAGAGATGTGGAGGAATACCGGTGGCGAAGGCGGCCCCCTGGACAAAGACTGACGCTC")

DS2.Proteus.plot <-ggplot(DS2.Proteus, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "DS2.Proteus abundance", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")

DESeq2_cont_DS2 <-ggarrange(DS2.Staphylococcus.plot, DS2.Proteus.plot, ncol = 3, nrow = 2)

# Dose 3

DS3.filt.cont <- DS3.filt %>%
transform_sample_counts(function(x) {x/sum(x)}) %>%
psmelt()

DS3.Shigella<- DS3.filt.cont %>% filter(OTU == "GCAAGCGTTAATCGGAATTACTGGGCGTAAAGCGCACGCAGGCGGTCAATTAAGTCAGATGTGAAAGCCCCGAGCTTAACTTGGGAATTGCATCTGAAACTGGTTGGCTAGAGTCTTGTAGAGGGGGGTAGAATTCCATGTGTAGCGGTGAAATGCGTAGAGATGTGGAGGAATACCGGTGGCGAAGGCGGCCCCCTGGACAAAGACTGACGCTC")

DS3.Shigella.plot <-ggplot(DS3.Shigella, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "DS3.Shigella abundance", color = "seroconversion")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")

DESeq2_cont_DS3 <-ggarrange(DS3.Shigella.plot, ncol = 3, nrow = 2)

```

```{r variable_correlation DS1}

multi.corr.DS1 <- psm.DS1 %>%
transform_sample_counts(function(x) {x/sum(x)}) %>%
preDA(min.samples = 10, min.reads = 0.01)
# filter_taxa(function(x) mean(x) > 0.01, TRUE)
# phyloseq_filter_prevalence(prev.trh = 0.1, abund.trh =0.01, abund.type = "mean", threshold_condition = "OR")

# View(data.frame(otu_table(multi.corr.DS1)))

multiple_correlation <- variable_correlation(multi.corr.DS1, variables = "postvaccination_IgA_titer", classification = "Species", method = "pearson")

# write.table(multiple_correlation, file = "./DS1_multiple_correlation_Species.txt", sep = "\t")

# Genus vs. postvaccination markers
DS1.Genus_all <- multi.corr.DS1 %>%
tax_glom(taxrank = "Genus") %>%
psmelt()

```

```{r variable_correlation DS2}

multi.corr.DS2 <- psm.DS2 %>%
transform_sample_counts(function(x) {x/sum(x)}) %>%
preDA(min.samples = 10, min.reads = 0.01)
# filter_taxa(function(x) mean(x) > 0.01, TRUE)
# phyloseq_filter_prevalence(prev.trh = 0.1, abund.trh =0.01, abund.type = "mean", threshold_condition = "OR")

# View(data.frame(otu_table(multi.corr.DS2)))

multiple_correlation_DS2 <- variable_correlation(multi.corr.DS2, variables = "postvaccination_IgA_titer", classification = "Genus", method = "pearson")

# write.table(multiple_correlation_DS2, file = "./DS2_multiple_correlation_Species.txt", sep = "\t")

# Genus vs. postvaccination markers
DS2.Genus_all <- multi.corr.DS2 %>%
tax_glom(taxrank = "Genus") %>%
psmelt()

multi.cor.Streptococcus <- DS2.Genus_all %>% filter(Genus == "Streptococcus")

multi.cor.Streptococcus.plot <-ggplot(multi.cor.Streptococcus, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(method="pearson") +
  labs(y = "postvaccination_IgA_titer", x = "Streptococcus abundance")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")

p.adjust(0.0067, method = "bonferroni", n = 44)

```

```{r variable_correlation DS3}

multi.corr.DS3 <- psm.DS3 %>%
transform_sample_counts(function(x) {x/sum(x)}) %>%
preDA(min.samples = 10, min.reads = 0.01)
# filter_taxa(function(x) mean(x) > 0.01, TRUE)
# phyloseq_filter_prevalence(prev.trh = 0.1, abund.trh =0.01, abund.type = "mean", threshold_condition = "OR")

# View(data.frame(otu_table(multi.corr.DS3)))

multiple_correlation_DS3 <- variable_correlation(multi.corr.DS3, variables = "postvaccination_IgA_titer", classification = "Genus", method = "pearson")

# write.table(multiple_correlation_DS3, file = "./DS3_multiple_correlation_Genus.txt", sep = "\t")

# Genus vs. postvaccination markers
DS3.Genus_all <- multi.corr.DS3 %>%
tax_glom(taxrank = "Genus") %>%
psmelt()

# write.table(DS3.Genus_all, file = "./DS3.Genus_all.txt", sep = "\t")

Eggerthella <- DS3.Genus_all %>% filter(Genus == "Eggerthella")

Eggerthella.plot <-ggplot(Eggerthella, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "Eggerthella abundance")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")

# Species vs. postvaccination markers
DS3.Species_all <- multi.corr.DS3 %>%
tax_glom(taxrank = "Species") %>%
psmelt()

# write.table(DS3.Species_all, file = "./DS3.Species_all.txt", sep = "\t")


lenta <- DS3.Species_all %>% filter(Species == "lenta")

lenta.plot <-ggplot(lenta, aes(x=Abundance, y=postvaccination_IgA_titer)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(y = "postvaccination_IgA_titer", x = "Eggerthella lenta abundance")+
  theme(plot.title = element_text(hjust=0.5), legend.position = "right")+
  # ylim(0,50)+
  stat_cor(method="pearson")

p.adjust(0.00011, method = "bonferroni", n = 44)

```



